---
title: "Final Project"
output: html_document
author: "Tyler Oh"
date: "Dec, 9th, 2020"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(devtools)
library(corrplot)
library(MASS)
library(e1071)
```

# Read incidence.csv and death.csv, clear and merge the dataset.

```{r}
incd = read_csv("incd.csv")
death = read_csv("death.csv")
```

## Each row of the County column in the incidence rate dataset contains brackets and values in it.
## Some rows of the column Average Annual Count in incidence rate data has the text "or fewer" at the end of values.

```{r}
incd$County = gsub("\\(.*?)", "",incd$County)
incd$`Age-Adjusted Incidence Rate(?) - cases per 100,000` = gsub("\\ #","",incd$`Age-Adjusted Incidence Rate(?) - cases per 100,000`)
```

## Merge the two dataset, incd and death, by County and FIPS.

```{r}
cancer = merge(incd,death,by=c("County","FIPS"))
```

## We only need County, FIPS, Incidence Rate, Death Rate, Average Deaths per Year
## After selecting the certain columns, we will rename columns' name for convenience.

```{r}
cancer <- cancer %>% 
  dplyr::select(c(1:3,12,15)) %>%
  rename(c(incidence_rate = `Age-Adjusted Incidence Rate(?) - cases per 100,000`,
           death_rate = `Age-Adjusted Death Rate`,
           avg_annual_death = "Average Deaths per Year"))
```

## If we change columns' type from char to num, all speical characters such as *, **, _ will be automatically
## converted to NA value. Hence, we will get rid of all NA values using na.omit().

```{r}
cancer$incidence_rate = as.numeric(cancer$incidence_rate)
summary(cancer)

cancer = na.omit(cancer)
summary(cancer)
```

## Dataset can be used here that has from census gov containing variables
## USA county data is https://github.com/Deleetdk/USA.county.data

## We will select some variables that relates to educational background, job occupation, health by County (FIPS).

```{r}
US_county = USA.county.data::USA_county_data
str(US_county,list.len=ncol(US_county))
```

# Choosing variables that seem to be reasonably related to the death rate from US County dateset.

## Educational background
Less.Than.High.School       
At.Least.High.School.Diploma     
At.Least.Bachelor.s.Degree         
Graduate.Degree
School.enrollment

## Poverty
Adults.65.and.Older.Living.in.Poverty  
Child.Poverty.living.in.families.below.the.poverty.line 
Poverty.Rate.below.federal.poverty.threshold

## Job occupation
Management.professional.and.related.occupations     
Service.occupations                                     
Sales.and.office.occupations                              
Construction.extraction.maintenance.and.repair.occupations: num 
Production.transportation.and.material.moving.occupations 

## Health
Poor.physical.health.days
Poor.mental.health.days
Adult.smoking
Adult.obesity

## Delete rows contating NA values and merge it with the original data by FIPS

```{r}
response_variables <- dplyr::select(US_county, c(1, 72:76, 85, 88, 90:93,95:96, 109:110, 114:115)) %>%
  na.omit(response_variables) %>%
  rename(c(FIPS = "fips"))
str(response_variables)
```

# Checking correlation among the chosen variables

## The variables in each of the four categories, Educational background, Poverty, Job occupation, Health should not be highly correlated.
## Therefore, we will find the correlation of each variables in same categories and delete one of the two variables that has high correlation.

## Educational Background:

## The plot shows high correlation between At.Least.High.School.Diploma and Less.Than.High.School and
## between Graduate.Degree and At.Least.Bachelor.s.Degree.
## It is fair to delete variables At.Least.High.School.Diploma and Graduate.Degree.

```{r}
education = dplyr::select(response_variables,c(2:6))
cor_educ = cor(education)
corrplot(method = "number", cor_educ)
```

## Poverty:

## We will remove Poverty.Rate.below.federal.poverty.threshold to avoid ~

```{r}
poverty = dplyr::select(response_variables,c(7:9))
cor_pov = cor(poverty)
corrplot(method = "number", cor_pov, tl.cex = 0.5)
```

## Job occupation:

## The plot shows no high corrlation among the job occupation variables.

```{r}
job = dplyr::select(response_variables,c(10:14))
cor_job = cor(job)
corrplot(cor_job, method = "number", tl.cex = 0.5)
```

# It seems there are potential high correlation between Poor.physical.health.days with Poor.mental.health.days
# However, we will keep all variables and do something with stepwise regression if it occurs problem.

```{r}
health = dplyr::select(response_variables,c(15:18))
cor_health = cor(health)
corrplot(cor_health, method = "number")
```

## -Graduate.Degree
## -At.Least.High.School.Diploma
## -Poverty.Rate.below.federal.poverty.threshold

```{r}
response_variables = response_variables[-c(3,5,8)]
```

## Merge the cancer data with chosen response variables by FIPS

```{r}
model = merge(cancer, response_variables, by = c("FIPS"))
str(model)
```

## We are going to select significant variables using forward selection, backward elimination and stepwise regression.

```{r}
y = model$death_rate
x1 = model$incidence_rate
x2 = model$avg_annual_death
x3 = model$Less.Than.High.School
x4 = model$At.Least.Bachelor.s.Degree
x5 = model$School.Enrollment
x6 = model$Adults.65.and.Older.Living.in.Poverty
x7 = model$Child.Poverty.living.in.families.below.the.poverty.line
x8 = model$Management.professional.and.related.occupations
x9 = model$Service.occupations
x10 = model$Sales.and.office.occupations
x11 = model$Construction.extraction.maintenance.and.repair.occupations
x12 = model$Production.transportation.and.material.moving.occupations
x13 = model$Poor.physical.health.days 
x14 = model$Poor.mental.health.days
x15 = model$Adult.smoking
x16 = model$Adult.obesity
```

# Histogram and boxplot of each variables

```{r}
par(mfrow=c(2,3))
hist(x1, right=FALSE)
hist(x2, right=FALSE)
hist(x3, right=FALSE)
hist(x4, right=FALSE)
hist(x5, right=FALSE)
hist(x6, right=FALSE)
hist(x7, right=FALSE)
hist(x8, right=FALSE)
hist(x9, right=FALSE)
hist(x10, right=FALSE)
hist(x11, right=FALSE)
hist(x12, right=FALSE)
hist(x13, right=FALSE)
hist(x14, right=FALSE)
hist(x15, right=FALSE)
hist(x16, right=FALSE)

par(mfrow=c(2,3))
boxplot(x1,xlab="Incidence Rate")
boxplot(x2,xlab="Average death per year")
boxplot(x3,xlab="Less than high school")
boxplot(x4,xlab="At least Bachelor degree")
boxplot(x5,xlab="School enrollment")
boxplot(x6,xlab="65 and older age living in poverty")
boxplot(x7,xlab="Child below the poverty line")
boxplot(x8,xlab="Management or professional occuptaions")
boxplot(x9,xlab="Service occupations")
boxplot(x10,xlab="Sales and office occupations")
boxplot(x11,xlab="Construction and repair occupations")
boxplot(x12,xlab="Production and transportation occupations")
boxplot(x13,xlab="Poor physical health days")
boxplot(x14,xlab="Poor mental health days")
boxplot(x15,xlab="Adult Smoking")
boxplot(x16,xlab="Adult Obesity")
```

```{r}
selected_data = as.data.frame(cbind(y,x1,x2,x3,x4,x5,x6,x7,x8,x9,x10,x11,x12,x13,x14,x15,x16))
```

# Compare residuals vs regression

```{r}
par(mfrow=c(2,2))
hist(log(x2), right=FALSE)
boxplot(log(x2),xlab="Average death per year")

Std_Res = rstandard(temp_death_fit)
plot(x2, Std_Res, ylab='Standardized Residuals', xlab='Average annual death', main='Average annual death (x2)', pch=10)
abline(h=0)
plot(log(x2), Std_Res, ylab='Standardized Residuals', xlab='Average annual death', main='Average annual death log(x2)', pch=10)
abline(h=0)
```

# Taking Logarithm.

```{r}
par(mfrow=c(1,2))
hist(log(x2), right=FALSE)
boxplot(log(x2),xlab="Average death per year")
hist(sqrt(x2), right=FALSE)
boxplot(sqrt(x2),xlab="Average death per year")
```

## Set null and full model for variable selection.

```{r}
fit_null = lm(y~1, data=selected_data)
fit_full = lm(y~x1+log(x2)+x3+x4+x5+x6+x7+x8+x9+x10+x11+x12+x13+x14+x15+x16, data=selected_data)
```

## Variable selection
## It seems all ways have almost same AIC

```{r}
forward=step(fit_null, data=selection_data, scope=list(lower=fit_null, upper=fit_full), direction="forward")
backward=step(fit_full, data=selection_data, direction="backward")
stepwise=step(fit_null, data=selection_data, scope=list(upper=fit_full), direction="both")
```

# It seems insignificant variables are still exist.

```{r}
summary(forward)
summary(backward)
summary(stepwise)

summary(update(stepwise,. ~.))
summary(update(stepwise,. ~. -x5))
summary(update(stepwise,. ~. -x5 -x14))

new_model = update(stepwise,. ~. -x5 -x14)
```

# Nomarlity test

```{r}
par(mfrow=c(1,3))
Res_death = resid(temp_death_fit)
Std_death = rstandard(temp_death_fit)
Stu_death = rstudent(temp_death_fit)
 
probplot(Res_death, qnorm, xlab='Residuals', ylab='Percent')
probplot(Std_death, qnorm, xlab='Standardized Residuals', ylab='Percent')
probplot(Stu_death, qnorm, xlab='Studentized Residuals', ylab='Percent')

par(mfrow=c(2,2))
plot(new_model)
```

```{r}
inf_model = influence.measures(new_model)
inf_mat = data.frame(inf_model$infmat)

which(inf_mat$cook.d > 0.025)
```


# Multicollinearity Check

## Since all values less than 10.

```{r}
vif(stepwise)
vif(new_model)
```

# Final model

```{r}
summary(new_model)
```


# Conclusion

# The best fitted model is y-hat = 1.0362 + 0.5781*x1 - 0.4369*log(x2) + 0.0664*x4 0.0465*x7 - 0.2175*x9 - 0.1316*x10 + 0.105859*x11 + 1.0494*x13 
# + 20.7525*x15+25.5870*x16 with Standard Error (5.842), Adjusted R-squared (0.8285), F-statistic (1063), and p-value (< 2.2e-16).
# All intendents variables are significant with set alpha (0.05). According to the final model, approximately 83% of the variability of death rate is explained through this regression.

# Corrplot
## The death rate is related to with factors in the order incidence rate, adult smoking rate and poor physical health days.
## Contrary to my expectations, there was no strong correlation between occupations or educational background with cancer incidence and mortality. ## However, we can notice that smoking and obesity contribute to invasive cancer mortality.

```{r}
final_set = as.data.frame(cbind(y,x1,log(x2),x4,x7,x9,x10,x11,x13,x15,x16))
corrplot(cor(final_set), method="number")
```



